\subsection{Cauchy's Integral Formula}
Cauchy's Integral Formula is an important representation of holomorphic functions as integrals which illuminates many important properties.
\begin{thm}[Cauchy's Integral Formula]
    Suppose $f$ is holomorphic on open set $\O$ which contains contour $T$ and its interior $I$. Then
    \[
        f(z) = \frac{1}{2\pi i}\int_T \frac{f(\z)}{\z - z}d\z
    \]
    for any $z \in I$.
\end{thm}
\begin{proof}
    Let $C_r$ denote the circle with radius $r$ centered at $z$. By the corridor argument, we have
    \[
        \frac{1}{2\pi i}\int_T \frac{f(\z)}{\z - z}d\z
        =
        \lim_{r \rightarrow 0} \frac{1}{2\pi i}\int_{C_r} \frac{f(\z)}{\z - z}d\z.
    \]
    Some standard computation shows the RHS is $f(z)$.
\end{proof}
There are two important consequences that immediately follow from Cauchy's Integral Formula.
\begin{thm}
    \label{holo-smooth}
    Holomorphic functions are infinitely differentiable.
\end{thm}
\begin{proof}
    Apply the difference quotient to Cauchy's Integral Formula to get
    \[
        f^{(n)} = \frac{n!}{2\pi i}\int_T \frac{f(\z)}{(\z-z)^{n+1}}d\z.
    \]
\end{proof}
\begin{remark}
    The formulas in the proof of Theorem \ref{holo-smooth} and Cauchy's Integral Formula are collectively known as the Cauchy Integral Formulas.
\end{remark}
\begin{thm}[Power Series Representation]
    Suppose $f$ is holomorphic on open set $\O$ which the circle contour $C$ centered at $z_0$ and its interior $I$. Then $f$ has a power series expansion at $z_0$ for all $z \in I$.
\end{thm}
\begin{details}{Proof overview}
    The idea is to convert Cauchy's Integral Formula into a geometric series. We use a clever trick to do this:
    \[
        \frac{1}{\z-z} = \frac{1}{\z -z_0} \frac{1}{1 - \frac{z-z_0}{\z-z_0}} = \frac{1}{\z-z_0}\sum_{n=0}^\i \Big( \frac{z-z_0}{\z-z_0} \Big)^n.
    \]
    We then plug this into Cauchy's Integral Formula.
\end{details}
\begin{remark}
    The reason we don't use an arbitrary contour here is a technical reason. A circular contour guarantees
    \[
        \Big| \frac{z - z_0}{\z - z_0} \Big| < 1
    \]
    Thus we can turn the fraction into a geometric series.
\end{remark}
