\section{Setting Up}

\subsection{Generalizing}
The quantum Hamiltonian of the wave equation for an electron in a hydrogen atom is:
\[
    \hat{H} = - \frac{\hbar^2}{2m}\Delta - \frac{Q^2}{|x|}
\]
where $m$ is the mass of the election, $x$ is the (3D) position of the election with respect to a fixed nucleus, and $Q$ is the charge of the electon.

We want to solve the corresponding wave equation. However, it happens that the method of solution in this system is virtually the same as in a general three dimensional quantum system with radial potential. Thus in this set of notes, we choose to generalize and solve the latter system.

The quantum Hamiltonian of the wave equation in a three-dimensional quantum system with radial potential is:
\[
    \hat{H} = - \frac{\hbar^2}{2m}\Delta + V
\]
where $V$ is any radial function in $\bR^3$. 

\subsection{}
As always, to solve the wave equation, we first find the eigenvectors of $\hat{H}$. This will be our next goal.

\subsection{Exploiting Symmetry}
The rotational symmetry of $\hat{H}$ implies that if we rotate a vector $\psi \in L^2(\bR^3)$, apply $\hat{H}$, and then rotate the resulting vector back, the final vector should be the same vector obtained by applying $\hat{H}$ to $\psi$.

More formally, consider the natural action of the rotation group $\SO(3)$ on $L^2(\bR^3)$ given by:
\[
    \psi(x) \mapsto \psi(R^{-1}x), \quad R \in \SO(3).
\]
This gives rise to a infinite dimensional unitary representation of $\SO(3)$:
\[
    (\Pi(R)\psi)(x) = \psi(R^{-1}x).
\]
(As mentioned in a previous set of notes, infinite dimensional representations require extra technical conditions to check, which we will not do here). Our discussion in the beginning can be formally written as:
\[
    \Pi(R)^{-1}\hat{H}\Pi(R)\psi = \hat{H}\psi.
\]
This implies $\hat{H}\Pi(R) = \Pi(R)\hat{H}$. That is, $\hat{H}$ commutes with all the rotation operators. Linear algebra tells us this implies the eigenspaces of $\hat{H}$ are invariant under $\Pi(R)$ for all $R \in \SO(3)$. In other words, the eigenspaces $\hat{H}$ induce representations of $\SO(3)$.

\subsection{(Optional) Angular Momentum Operators}
Let $\pi$ be the Lie algebra representation $\so(3) \rightarrow \gl(L^2(\bR^3))$ induced by $\Pi$. Let $F_1, F_2, F_3$ be the usual basis of $\so(3)$ (see the set of notes of $\SO(3), \so(3)$). Then it happens that
\[
    i\hbar\pi(F_j)
\]
are the componentwise angular momentum operators. More specifically
\[
    \begin{split}
        i\hbar\pi(F_1) &= X_2P_3 - X_3P_2\\
        i\hbar\pi(F_2) &= X_3P_1 - X_1P_3\\
        i\hbar\pi(F_3) &= X_1P_2 - X_2P_1\\
    \end{split}
\]
where $X_i, P_i$ are the position and momentum operators in each component. Considering the relationship between Lie groups and Lie algebras, this implies that the angular momentum operators are the ``infinitesimal" generators of the rotation group operators.

\subsubsection{Aside}
Attentive readers may have noticed that $\GL(L^2(\bR^3)$ is not a Lie group because it is infinite-dimensional, and by the same reason, nor is $\gl(L^2(\bR^3))$ a Lie algebra. How can we then define the Lie algebra representation $\pi$ induced by $\Pi$? The definition we have been using is done in the finite dimensional case, where $\GL$ is a Lie group and $\gl$ is a Lie algebra, so we were able to consider Lie algebra homomorphisms induced by Lie group homomorphisms.

The answer lies in a slight workaround. Let $\mathfrak g$ be the Lie algebra corresponding to $G$. Recall
\[
    \mathfrak g = \{X \in M_n(\bC) : e^{tX} \in G \text{ for all } t \in \bR\}.
\]
If $\Pi'$ is a Lie group homomorphism from $G$, the induced Lie algebra homormophism $\pi'$ can alternatively be defined as the map that satisfies
\[
    \Pi'(e^{tX}) = e^{t\pi'(X)}
\]
This is an equivalent definition to the one we have been using. However, this definition can be used to define induced homomorphisms when the target is no longer a Lie group. Thus we use it to define $\pi$:
\[
    \Pi(e^{tX}) = e^{t\pi(X)}.
\]
This definition also allows us to more rigorously say that the angular momentum operators are the ``infinitesimal" generators of the rotation group operators (previously we were intuitively assuming $\GL(L^2(\bR^3))$ was a Lie group and $\gl(L^2(\bR^3))$ was the corresponding Lie algebra).
