\section{Lecture 3-3-2020}
\textbf{Some observations.} Working in a lattice space allows us to use a few tricks:
\begin{itemize}
    \item We can decompose $x = x^+ - x^-$, so we need only study positive elements. We have already seen this many times.
    \item Given $x$, we can create elements like $x \vee 1$, $x \wedge 1$, etc. More generally, we can ``combine" multiple elements. This is a powerful property for construction (c.f. Stone Weierstrass).
\end{itemize}
The following toy example will show the power of the second trick. Then we will use the second trick to prove partitions of unity.

\begin{thm}[A Toy Example]
Suppose $\mu, \nu$ are real measures such that $\mu \wedge \nu = 0$. Then $\mu$ and $\nu$ are mutually singular.
\end{thm}

\begin{details}{Proof gist}
If not, construct a strictly positive measure smaller than both.
\end{details}

\begin{details}{More detail}
Afer applying Lebesgue Decomposition and then R-N, we get
\[
    \int_E h d\mu = \nu_{ac}(E).
\]
If $h \neq 0$ a.e., consider the strictly positive measure induced by $h \wedge 1$.
\end{details}

\begin{thm}[Partitions of Unity for LCH Spaces]
If $X$ is LCH, then for any compact $C \subset X$ and open cover $\theta_i$ of $C$, there exists positive functions $f_i \in C_C(X)$ such that $\sum f_i = 1$ on $C$ and supp$(f_i) \subset \theta_i$.
\end{thm}
\begin{remark}
Support is defined as the closure of the carrier.
\end{remark}
\begin{details}{Proof gist}
Use LCH Urysohn's lemma (and friend) to construct a function
\[
    g = \sum g_i \geq 1.
\]
Then consider
\[
    \frac{g}{g \vee 1}.
\]
\end{details}

\begin{details}{Construction of first equation}
We first show there exists closed sets $B_i \subset \theta_i$ such that $C \subset \cup B_i$. (Need LCH Urysohn's friend). Then use LCH Urysohn's lemma to define $g_i(x) = 1$ for $x \in B_i$ and supp$(g_i) \subset \theta_i$.
\begin{itemize}
    \item \textbf{LCH Urysohn's friend (used to prove LCH Urysohn).} $X$ LCH, subset $C$ compact, $U$ open, $C \subset U$. Then there exists open $V$, $\bar{V}$ compact, such that $C \subset V \subset \bar{V} \subset U$.
    \item \textbf{LCH Urysohn's lemma.} If subset $C$ is compact and $C \subset \theta$ is open, then there exists a continuous $f: X \rightarrow [0,1]$ with $f(x) = 1$ for $x \in C$ and $f(x) = 0$ for $x \not \in \theta$.
\end{itemize}
Sorry, I come up with weird names.
\end{details}
